---
title: "Linear Model Analysis"
output: html_notebook
---

Our goal is to simply model the data, with and without the PV data.  To do this,
I am using two models: a linear model (LM) and a generalised linear model (GLM).

This code contains two main components for each - an EXPLANATORY model and a
PREDICTIVE model (roll-forward).

Libraries
```{r include=FALSE}
library(data.table)
library(ggplot2)
library(plotly)
library(animation)
library(lubridate)
library(stats)
library(beepr)
library(datawizard)
```

Global variables
```{r}
startDT <- as.POSIXct("2018-03-07 00:00:00", format="%Y-%m-%d %H:%M:%S",tz = "Etc/GMT-10")
stopDT  <- as.POSIXct("2021-01-31 23:30:00", format="%Y-%m-%d %H:%M:%S",tz = "Etc/GMT-10")
sampleTimes <- seq(from=startDT, to=stopDT, by = "30 min") # 
```

Functions
```{r}
# resamples and interpolates data at new time points
resampleData <- function(x, y, newx){
  fit <- approxfun(x,y, rule = 2:2) # linear
  resampled = fit(newx)
  return(resampled)
}

# uniformly resamples a particular set of state data
cleanState <- function(dataState){
  newState <- data.table(
    DATETIME = sampleTimes,
    TOTALDEMAND = resampleData(dataState$DATETIME, dataState$TOTALDEMAND,sampleTimes),
    POWER = resampleData(dataState$DATETIME, dataState$POWER,sampleTimes),
    TEMPERATURE = resampleData(dataState$DATETIME, dataState$TEMPERATURE,sampleTimes))
  return(newState)
}

# computes the MAPE
mape <- function(real, pred){
  return(100 * mean(abs((real - pred)/real)))
}

# plot the model over a particular range
plotModel <- function(model, xrange, times){
  ydata <- model$model$DEMAND
  plot(times[xrange],ydata[xrange])
  lines(times[xrange],model$fitted.values[xrange],col = "red",lwd=3)
  mape(ydata , model$fitted.values)
}

# use a polynomial to predict ONE position further.
polyPred <- function(data,k){
  x <- seq(1:length(data))
  model <- lm(data ~ poly(x,k))
  xnew <- data.frame(x = (length(data)+1))
  y <- predict(model, newdata=xnew)
}

```


Load and clean the raw data
```{r}
allTheData <- read.csv("dataset_v1_0.csv")
allTheData <- allTheData[complete.cases(allTheData), ]
allTheData$DATETIME <- as.POSIXct(allTheData$DATETIME, format="%Y-%m-%d %H:%M:%S",tz = "Etc/GMT-10")
allTheData <- subset(allTheData, DATETIME >= startDT)
allTheData <- subset(allTheData, DATETIME < stopDT)

toKeep <-c("DATETIME","TOTALDEMAND","POWER","TEMPERATURE")
dataNSW <- subset(allTheData, STATE=="NSW")[toKeep]
dataQLD <- subset(allTheData, STATE=="QLD")[toKeep]
dataVIC <- subset(allTheData, STATE=="VIC")[toKeep]
dataSA <- subset(allTheData, STATE=="SA")[toKeep]
rm(allTheData)

dataNSW <- cleanState(dataNSW)
dataQLD <- cleanState(dataQLD)
dataVIC <- cleanState(dataQLD)
dataSA <- cleanState(dataSA)
```

Create categorical factors
```{r}
# public holidays categorical variable
pub.hols.NSW <- tsibble::holiday_aus(2018:2022,state="NSW")
pub.hols.QLD <- tsibble::holiday_aus(2018:2022,state="QLD")
pub.hols.VIC <- tsibble::holiday_aus(2018:2022,state="VIC")
pub.hols.SA <-  tsibble::holiday_aus(2018:2022,state="SA")

pub.hols.NSW <- date(as.POSIXct(pub.hols.NSW$date, format="%Y-%M-%D",tz = "Etc/GMT-10"))
pub.hols.QLD <- date(as.POSIXct(pub.hols.QLD$date, format="%Y-%M-%D",tz = "Etc/GMT-10"))
pub.hols.VIC <- date(as.POSIXct(pub.hols.VIC$date, format="%Y-%M-%D",tz = "Etc/GMT-10"))
pub.hols.SA  <- date(as.POSIXct(pub.hols.SA$date,  format="%Y-%M-%D",tz = "Etc/GMT-10"))

# seasons categorical variable
# summer = 1, autumn = 2, winter = 3, fall = 4
# NOTE: this is NOT based on meterological seasons, but rather summer being the
# hottest weeks and winter being the coldest weeks
seasons <- rep(1,55)
seasons[09:21] <- 2
seasons[22:34] <- 3
seasons[35:47] <- 4
```

Create Linear Model Data Table
```{r}
N <- 1488 # number of time slots in January 2021.
L <- 49488 # length of the training set

state = "NSW"
if (state=="NSW"){
  data <- dataNSW
  pub.hols <- pub.hols.NSW
}

# create data table holding all the model data
genSupply <- data.table(DEMAND = data$TOTALDEMAND,
                        DATE1 = standardize(1:50976),
                        DATE2 = standardize((1:50976))^2,
                        DATE3 = standardize((1:50976))^3,
                        TEMPERATURE1 = standardize(data$TEMPERATURE),
                        TEMPERATURE2 = standardize(data$TEMPERATURE^2),
                        TEMPERATURE3 = standardize(data$TEMPERATURE^3),
                        dayindex = as.factor(rep(1:48,1062)),
                        dayName = as.factor(strftime(sampleTimes,format = "%w")),
                        season = as.factor(seasons[(1+as.integer(strftime(sampleTimes,format="%W")))]),
#                        weekNum = as.factor(as.integer(strftime(sampleTimes,format="%W"))),
                        pubHoliday = as.integer(date(sampleTimes) %in% (pub.hols)))
actDemand <- genSupply
actDemand$DEMAND <- actDemand$DEMAND + data$POWER

#create emptly list to hold results for WITH and WITHOUT PV data
genSupplyForecast <- rep(0,N) #results for not considering PV data
actDemandForecast <- rep(0,N) #results for when considering PV data
```


Linear Model Exploration (X for exploratory)
```{r}
genSupplyX <- lm(log(DEMAND) ~ 1 + ., data = genSupply)
summary(genSupplyX)
# REMEMBER: when looking at these terms, they need to be considered in terms
# of EXPONENTIAL values.
```


Linear Model (LM) Roll Forward Analysis
```{r}
# a short note
# I used this same code to explore the optimal weights, the optimal polymer 
# prediction function, and the optimal number of previous elements to use for
# prediction.  Turns out, a linear model using the last 10 time slots.


weights = rev(exp(-(0:(L-1))/64)) #compute weights - effectively only three days
j=10
for (i in 1:N){
    # generate the LM for Generated Supply (does not consider PV data)
    suppressWarnings(genSupplyLM <- lm(log(DEMAND) ~ 1 + ., data = genSupply[i:(i+L-1),], weights = weights))
    genSupplyData <-genSupply[i+L,]
    genSupplyData$DEMAND <-0 #this is to ensure that we simply don't copy the result
    genSupplyData$TEMPERATURE1 <- polyPred(genSupply$TEMPERATURE1[(i+L-1-j):(i+L-1)],1)
    genSupplyData$TEMPERATURE2 <- polyPred(genSupply$TEMPERATURE2[(i+L-1-j):(i+L-1)],1)
    genSupplyData$TEMPERATURE3 <- polyPred(genSupply$TEMPERATURE3[(i+L-1-j):(i+L-1)],1)
    suppressWarnings(genSupplyForecast[i] <- predict(genSupplyLM,newdata=genSupplyData))

    suppressWarnings(actDemandLM <- lm(log(DEMAND) ~ 1 + ., 
                                       data = actDemand[i:(i+L-1),],weights = weights))
    actDemandData <-actDemand[i+L,]
    actDemandData$DEMAND <-0 #this is to ensure that we simply don't copy the result
    actDemandData$TEMPERATURE1 <- polyPred(actDemand$TEMPERATURE1[(i+L-1-j):(i+L-1)],1)
    actDemandData$TEMPERATURE2 <- polyPred(actDemand$TEMPERATURE2[(i+L-1-j):(i+L-1)],1)
    actDemandData$TEMPERATURE3 <- polyPred(actDemand$TEMPERATURE3[(i+L-1-j):(i+L-1)],1)
    suppressWarnings(actDemandForecast[i] <- predict(actDemandLM,newdata=actDemandData))
    if (!(i %% 10)){print(i)}
}


plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Generated Supply (no PV included)", xlab = "Date (2021)", ylab = "Power",
     xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],exp(genSupplyForecast[1:N]) ,
      col='red', lwd=2)

plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Actual Demand (with PV included)",xlab = "Date (2021)", ylab = "Power",
    xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],exp(actDemandForecast[1:N])-data$POWER[(L+1):(L+N)],
      col='green', lwd=2)

noPVoutput <- data.table(DATETIME = sampleTimes[(1+L):(N+L)], TOTALDEMAND = exp(genSupplyForecast[1:N]))
withPVoutput <- data.table(DATETIME = sampleTimes[(1+L):(N+L)], TOTALDEMAND = exp(actDemandForecast[1:N])-data$POWER[(L+1):(L+N)])

write.csv(noPVoutput, paste(state," - LM - without future info - without PV.csv"), row.names=FALSE)
write.csv(withPVoutput, paste(state, " - LM - without future info - with PV.csv"), row.names=FALSE)

print(mape(data$TOTALDEMAND[(1+L):(N+L)],exp(genSupplyForecast[1:N])))
print(mape(data$TOTALDEMAND[(1+L):(N+L)] + data$POWER[(1+L):(N+L)],exp(actDemandForecast[1:N])))

```



Explore weights and prediction model
Here, the variable 'j' can be used to explore the effects of weightings
```{r}

for (j in 4:9){
  weights = rev(exp(-(0:(L-1))/64)) #compute
#  weights = rev(exp(-(0:(L-1))/(2^j)) #compute

  for (i in 1:N){
    # generate the LM for Generated Supply (does not consider PV data)
    suppressWarnings(genSupplyLM <- lm(log(DEMAND) ~ 1 + ., 
                                       data = genSupply[i:(i+L-1),], weights = weights))
    genSupplyData <-genSupply[i+L,]
    genSupplyData$DEMAND <-0 #this is to ensure that we simply don't copy the result
#    genSupplyData$TEMPERATURE1 <- polyPred(genSupply$TEMPERATURE1[(i+L-1-j):(i+L-1)],0)
    genSupplyData$TEMPERATURE1 <- polyPred(genSupply$TEMPERATURE1[(i+L-1-10):(i+L-1)],1)
    genSupplyData$TEMPERATURE2 <- polyPred(genSupply$TEMPERATURE2[(i+L-1-10):(i+L-1)],1)
    genSupplyData$TEMPERATURE3 <- polyPred(genSupply$TEMPERATURE3[(i+L-1-10):(i+L-1)],1)
    suppressWarnings(genSupplyForecast[i] <- predict(genSupplyLM,newdata=genSupplyData))

    suppressWarnings(actDemandLM <- lm(log(DEMAND) ~ 1 + ., 
                                       data = actDemand[i:(i+L-1),],weights = weights))
    actDemandData <-actDemand[i+L,]
    actDemandData$DEMAND <-0 #this is to ensure that we simply don't copy the result
    actDemandData$TEMPERATURE1 <- mean(actDemand$TEMPERATURE1[(i+L-1-j):(i+L-1)])
    actDemandData$TEMPERATURE2 <- mean(actDemand$TEMPERATURE2[(i+L-1-j):(i+L-1)])
    actDemandData$TEMPERATURE3 <- mean(actDemand$TEMPERATURE3[(i+L-1-j):(i+L-1)])
    suppressWarnings(actDemandForecast[i] <- predict(actDemandLM,newdata=actDemandData))
  }
  print(j)
  print(mape(data$TOTALDEMAND[(1+L):(N+L)],exp(genSupplyForecast[1:N])))
  print(mape(data$TOTALDEMAND[(1+L):(N+L)] + data$POWER[(1+L):(N+L)],exp(actDemandForecast[1:N])))
}
```


GLM models
```{r}
N <- 1488
L <- 49488 # length of the training set

state = "NSW"
if (state=="NSW"){
  data <- dataNSW
  pub.hols <- pub.hols.NSW
}

genSupplyForecast <- rep(0,N) #results for not considering PV data
weights = rev(exp(-(0:(L-1))/640)) #compute
j=10
for (i in 1:N){
  # generate the GLM for Generated Supply (does not consider PV data)
#  genSupplyGLM <- glm(log(DEMAND) ~ ., data = genSupply[i:(i+L-1),], weights = weights)
  genSupplyGLM <- glm(DEMAND ~
    DATE1*(dayindex+dayName+season+pubHoliday)+
    DATE2*(dayindex+dayName+season+pubHoliday)+
    DATE3*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE1*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE2*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE3*(dayindex+dayName+season+pubHoliday),
    data = genSupply[i:(i+L-1),], weights = weights)

  genSupplyData <-genSupply[i+L,]
  genSupplyData$DEMAND <-0 #this is to ensure that we simply don't copy the result
  genSupplyData$TEMPERATURE1 <- polyPred(genSupply$TEMPERATURE1[(i+L-1-j):(i+L-1)],1)
  genSupplyData$TEMPERATURE2 <- polyPred(genSupply$TEMPERATURE2[(i+L-1-j):(i+L-1)],1)
  genSupplyData$TEMPERATURE3 <- polyPred(genSupply$TEMPERATURE3[(i+L-1-j):(i+L-1)],1)
  genSupplyForecast[i] <- predict(genSupplyGLM,newdata=genSupplyData)

#  actDemandGLM <- glm(log(DEMAND) ~ ., data = actDemand[i:(i+L-1),],weights = weights)
  actDemandGLM <- glm(DEMAND ~
    DATE1*(dayindex+dayName+season+pubHoliday)+
    DATE2*(dayindex+dayName+season+pubHoliday)+
    DATE3*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE1*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE2*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE3*(dayindex+dayName+season+pubHoliday),
    data = actDemand[i:(i+L-1),], weights = weights)
  
  actDemandData <-actDemand[i+L,]
  actDemandData$DEMAND <-0 #this is to ensure that we simply don't copy the result
  actDemandData$TEMPERATURE1 <- polyPred(actDemand$TEMPERATURE1[(i+L-1-j):(i+L-1)],1)
  actDemandData$TEMPERATURE2 <- polyPred(actDemand$TEMPERATURE2[(i+L-1-j):(i+L-1)],1)
  actDemandData$TEMPERATURE3 <- polyPred(actDemand$TEMPERATURE3[(i+L-1-j):(i+L-1)],1)
  actDemandForecast[i] <- predict(actDemandGLM,newdata=actDemandData)
  print(i)
}

plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Generated Supply (no PV included)", xlab = "Date (2021)", ylab = "Power",
     xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],(genSupplyForecast[1:N]) ,
      col='red', lwd=2)

plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Actual Demand (with PV included)",xlab = "Date (2021)", ylab = "Power",
    xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],(actDemandForecast[1:N])-data$POWER[(L+1):(L+N)],
      col='green', lwd=2)
  
noPVoutput <- data.table(DATETIME = sampleTimes[(1+L):(N+L)], TOTALDEMAND = (genSupplyForecast[1:N]))
withPVoutput <- data.table(DATETIME = sampleTimes[(1+L):(N+L)], TOTALDEMAND = (actDemandForecast[1:N])-data$POWER[(L+1):(L+N)])

write.csv(noPVoutput, paste(state," - GAM - without future info - without PV.csv"), row.names=FALSE)
write.csv(withPVoutput, paste(state, " - GAM - without future info - with PV.csv"), row.names=FALSE)

print(mape(data$TOTALDEMAND[(1+L):(N+L)],(genSupplyForecast[1:N])))
print(mape(data$TOTALDEMAND[(1+L):(N+L)] + data$POWER[(1+L):(N+L)],(actDemandForecast[1:N])))
```


GAM model (UNFINISHED)
```{r}


  genSupplyGAM1 <- mgcv::gam(log(DEMAND) ~
    s(DATE1,pubHoliday,k=100)+
    s(DATE2,pubHoliday,k=100)+
    s(DATE3,pubHoliday,k=100)+
    s(TEMPERATURE1,pubHoliday,k=100)+
    s(TEMPERATURE2,(pubHoliday),k=100) +
    s(TEMPERATURE3,(pubHoliday),k=100) +
    season +dayName+ dayindex ,
    data = genSupply)
beep(1)
    

   +
    genSupply$season + genSupply$dayindex + genSupply$dayName


  genSupplyGAM2 <- mgcv::gam(
  genSupply$DEMAND ~ s(genSupply$DATE1,genSupply$DATE2,
  genSupply$TEMPERATURE1, genSupply$TEMPERATURE2, genSupply$TEMPERATURE3,
  genSupply$pubHoliday,
  k=100) + genSupply$season + genSupply$dayindex + genSupply$dayName,
  )

  genSupplyGLM <- glm(DEMAND ~
    DATE1*(dayindex+dayName+season+pubHoliday)+
    DATE2*(dayindex+dayName+season+pubHoliday)+
    DATE3*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE1*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE2*(dayindex+dayName+season+pubHoliday)+
    TEMPERATURE3*(dayindex+dayName+season+pubHoliday),
    data = genSupply[i:(i+L-1),], weights = weights)

  plot(genSupply$DEMAND[1:200])
lines(genSupplyGAM2$fitted.values[1:200])

beep(3)

suppressWarnings(
  actDemandGAM1 <- mgcv::gam(
  actDemand$DEMAND ~ s(actDemand$DATE1,actDemand$DATE2,
  actDemand$TEMPERATURE1, actDemand$TEMPERATURE2, actDemand$TEMPERATURE3,
  actDemand$pubHoliday,
  k=1000) + actDemand$season + actDemand$dayindex + actDemand$dayName,
  ))
beep(3)

plot(genSupply$DEMAND[1:200])
plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Generated Supply (no PV included)", xlab = "Date (2021)", ylab = "Power",
     xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],(genSupplyForecast[1:N]) ,
      col='red', lwd=2)

plot(sampleTimes[(1+L):(N+L)],data$TOTALDEMAND[(1+L):(N+L)],
     main = "NSW: Actual Demand (with PV included)",xlab = "Date (2021)", ylab = "Power",
    xlim = c(sampleTimes[L],sampleTimes[L+500]))
lines(sampleTimes[(1+L):(N+L)],(actDemandForecast[1:N])-data$POWER[(L+1):(L+N)],
      col='green', lwd=2)

```


